---
title: "practical-1"
author: "Sumit"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---
# M.Sc-2    
## Q-1 
### Create a new object x that stores the values (2,4,3,5,7,9).Run the following commands on R. Also give your comments.
```{r}
x<-c(2,4,3,5,7,9)
length(x)           #counts no. of elements in the vector.
sum(x)            #adds all the elements.
mean(x)          #it calculates the average.
min(x)          #smallest element of the vector
max(x)          #largest element of the vector
range(x)#it returns a vector containing the minimum and maximum of all the given arguments.
median(x)          #middle value
quantile(x)          #finding quantiles
summary(x)          #finding summary
sort(x)          #sort a vector into ascending or descending order.
rev(sort(x))          #it reverse the vector 
order(x)#it will sort the given numbers according to its index in the ascending order
x[order(x)]       #it gives the values of corresponding indexes
x[2]              #it returns value of 2nd index
x[c(2,4)]          #it returns value of 2nd and 4th index.
x[-2]               #it deletes the value at given index from vector
x[-c(1,3)]          #removes element from 1st and 3rd index 
log(x)              # finding the log value of vector x
log(x,base=10)      #finding the log value to base 10
log(x,base = 2)     #finding the log value to base 10
1+x                 #adding 1 to each element of x
2*x                 #multip1ying 2 to each element of x
log(1+x)            #finding log after adding 1
log(2*x)            # finding log after multiplying 2
sqrt(x)             #finding square root of the vector
x^2                 #calculating the square. 
x^0.32              #calculating the power 0.32 of x
sum(x^0.32)         # calculating the sum of x^0.32
sin(x)
cos(x)
tan(x)
170166719/31079     #Division 
170166719%%31079    #Remainder
170166719%/%31079   #Quotient
# (i) sort(x) Vs order(x)
# sort(x) means arranging the elements of vector in increasing order and order(x) means the index
#positions of this sorted values.
# (ii)sort(x) Vs x[order(x)]
# sort(x) means arranging the elements of vector in increasing order while x[order(x)] is giving
#the corresponding sub vector of x according to the index positions of order(x)

```
Q-2 Using rep() and seq() as needed,create the vectors
```{r}
a<-seq(0,4)
rep(a,each=5)
b<-seq(1,5)
rep(b,5)
```
Q-3 Using rep() and seq() as needed,create the vectors
```{r}
x<-c(1:5,2:6,3:7,4:8,5:9)
x

# OR
rep(0:4,each=5)+seq(1:5)

```
Q-4 Create and store a sequence of values from 5 to -11 that progress in steps of 0.3
```{r}
seq(5,-11,by=-0.3)
```
Q-5 The following are a sample of observations on incoming solar radiation at a green house.
```{r}
# a) Assign the data to the object called solar.radiation
solar.radiation=c( 11.1,10.6,6.3,8.8,10.7,11.2,8.9,12.2)
# b) Find the mean, median,variance of the radiation observations:
mean(solar.radiation)
median(solar.radiation)
var(solar.radiation)
# c) Add 10 to each observation of the solar.radiation and assign
#the result to sr10.find the mean,median and variance of sr10.
#which statistics change and by how much.
(sr10=solar.radiation+10)
mean(sr10)
median(sr10)
var(sr10)
# Conclusion: Mean and median increased by 10
# d) multiply each observation by -2 and assign the result to srm2.
# find the mean, median and variance of srm2. How do the statistics
# change Now.
(srm2=solar.radiation*(-2))
mean(srm2)
median(srm2)
var(srm2)
# Conclusion: mean is decreased by 9.975,median is decreased by 10.65
#and variance is increased by 10.575
# e)
# There are two formulas commonly used for the variance of a set of numbers .
# one uses the sample size n in the denominator and the other uses (n-1).
# Which formula does the var() function in R Use?
# i) Calculating variance using denominator 'n'
(y=solar.radiation-mean(solar.radiation))
(v1=sum(y^2)/length(solar.radiation))
# ii) Calculating variance using denominator 'n-1'
(v2=sum(y^2)/(length(solar.radiation)-1))
# iii) Calculating variance in 'R'
var(solar.radiation)
# conclusion: So the var() function in 'R' uses (n-1) as denominator.
```
Q_6:To convert the given temperatures (45,77,20,19,101,120,212) Fahrenheit to degree celsius.
```{r}
F=c(45,77,20,19,101,120,212) # temperatures in Fahrenheit
# converting to degree celsius
c=5*(F-32)/9
c
```
